---
title: "Channel Attribution"
author: "Wanyu(Crystal) Lee"
date: "9/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(knitr)
library(kableExtra)
library(ChannelAttribution)
```

## Dataset
The dataset is from [an article in Medium](https://towardsdatascience.com/marketing-channel-attribution-with-markov-chains-in-python-part-2-the-complete-walkthrough-733c65b23323). It tracked customer journeys from July 1, 2018, to July 31, 2018, and detailedly recorded when each ad showed up, which customer it reached, and whether the customer successfully converted. 

For example, for the first cookie, its journey includes four states: Instagram > Online Display > Online Display > Online Display. Unfortunately, the user did not convert during the observed period, and thus both conversion and conversion value are 0.

```{r}
dt = fread('attribution data.csv')

head(dt) %>% kable() %>%
           kable_styling(bootstrap_options = "striped",
                         full_width = F, 
                         position="center")
```

* **cookie:** used for identifying unique customers 
* **time**
* **interaction:** "impression" or "conversion". 
* **conversion:** 0 or 1
* **conversion_value:** how much values a consumer brings about after conversion
* **channel:** including "Facebook", "Instagram", "Online Display", "Online Video", and "Paid Search"



## Preprocessing

In order to apply the Markov Chain model, we have to transform our data and create a path variable like our previous example. Here I used `data.table` to deal with our data because it can process data much more quickly than `data.frame`.

```{r}
library(lubridate)
library(stringr)

dt = dt[order(cookie, time),time:=ymd_hms(time)][,id := seq_len(.N), by = cookie]

dt_wide = dcast(data = dt, formula = cookie ~ id, value.var = "channel")
dt_wide = dt_wide[, path:=do.call(paste,c(.SD, sep=' > ')), .SDcols=-1]
dt_wide = dt_wide[, path:=word(path, 1, sep = " > NA")]

conversion = dt[, .(conversion=sum(conversion), conversion_value=sum(conversion_value)), by=cookie]

setkey(conversion, cookie)
setkey(dt_wide, cookie)

dt_wide = merge(dt_wide, conversion)

head(dt_wide[, .(path, conversion, conversion_value)]) %>%
           kable() %>%
           kable_styling(bootstrap_options = "striped",
                         full_width = F, 
                         position="center")
```

## Markov Chain Modeling

R has a great package designed for channel attribution, called ChannelAttribution. It can be used to build models based on heuristic and markovian approaches, respectively. To evaluate the Markov Chain results, I simultaneously run heuristic models and regard them as base models.

```{r}
# Apply models
H = heuristic_models(Data = dt_wide, 
                     var_path = 'path', 
                     var_conv = 'conversion', 
                     var_value='conversion_value', 
                     sep='>')

M = markov_model(Data = dt_wide, 
                     var_path = 'path', 
                     var_conv = 'conversion', 
                     var_value='conversion_value', 
                     sep='>', 
                     order=1, 
                     out_more=TRUE)

M$result %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F, 
                position="center")
```


```{r}
H %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F, 
                position="center")
```


```{r}
M$transition_matrix %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F, 
                position="center")
```

```{r}
M$transition_matrix %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F, 
                position="center")
```

```{r}
M$removal_effects %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped",
                full_width = F, 
                position="center")
```

